Outline paper - Multiple imputation for statistical disclosure control: Creating synthetic datasets with *mice*

### 2020-11-20 mice_synthesizing/paper.Rmd
Add first part of the mice approach section.

- [x] What is synthetic data?
- [x] Why is it useful?
- [x] Why should it be easy to generate?
- [x] Parallel to missing data
- [x] If similar to missing data, then why not use `mice`?
- [ ] `mice` can only be exploited if the statistical properties are correct.
- [ ] In this paper we demonstrate a straightforward workflow for generating synthetic datasets and we investigate how to obtain valid inferences from synthesized data.

@@

Open science, including open data, has been marked as the future of science [@gewin_data_2016], and the advantages of publicly available research data are numerous [@molloy_open_2011; @walport_brest_sharing_2011]. Because collecting research data requires an enormous investment both in terms of time and monetary resources, openly accessible research data bears the potential of increasing the scientific returns. Additionally, the fact that public funds are used for data collection results in increased demands for the collected data. Nevertheless, the possibilities to share research data directly are often very limited due to restrictions with regard to privacy and confidentiality. Simply anonimizing the data is not enough to fulfill these requirements [@ohm_broken_2009]. Over the years, several other techniques have been used to increase the confidentiality of the data, such as categorizing continuous variables, top coding values above a certain threshold or adding random noise [@drechsler_synthetic_2011]. However, these methods may distort relationships between variables, thereby drastically reducing the data quality and the usability for further research.

@@

The `mice` package has been developed originally for multiple imputation of missing data, but can be easily extended to the synthetic data framework. It proceeds by drawing missing or synthetic values from the joint distribution of the observed data. This joint density is generally hard or impossible to derive, which can be overcome by employing fully conditional specification (FCS). Consider a $n \times k$ data matrix $Y$, with values to be replaced $Y^{old}$, the values to be kept $Y^{keep}$ and the values that are already replaced $Y^{new}$, under the assumption that there are no missing values in the observed data set. Instead of drawing from the joint distribution, we draw from $k$ univariate conditional densities for each variable. That is, we draw from the distribution $P\big{(}Y_j^{old} | Y^{keep}, Y^{new}, Y^{old}_{-j}, \theta \big{)}$, where the drawn values are conditional upon the values that are to be kept, the synthetic values that are drawn at an earlier iteration and the values that still are to be replaced, with $j = 1, 2, \dots, k$ and $\theta$ indicating the parameter specifying the distribution of $Y_j^{old}$. Thus, the synthetic values are generated variable by variable. This procedure is executed $i = 1, 2, \dots, m$ times, resulting in $\textbf{D} = \{D{(1)}, \dots, D^{(m)}\}$ synthetic datasets, with $D^{(i)} = (Y^{keep}, Y^{i, new})$.

@@

>>>>

- [x] What is synthetic data?
- [x] Why is it useful?
- [x] Why should it be easy to generate?
- [x] Parallel to missing data
- [x] If similar to missing data, then why not use `mice`?
- [x] `mice` can only be exploited if the statistical properties are correct.
- [x] In this paper we demonstrate a straightforward workflow for generating synthetic datasets and we investigate how to obtain valid inferences from synthesized data.

@@

Open science, including open data, has been marked as the future of science [@gewin_data_2016], and the advantages of publicly available research data are numerous [@molloy_open_2011; @walport_brest_sharing_2011]. Because collecting research data requires an enormous investment both in terms of time and monetary resources, openly accessible research data bears the potential of increasing the scientific returns. Additionally, the fact that public funds are used for data collection results in increased demands for the collected data. Nevertheless, the possibilities to share research data directly are often very limited due to restrictions with regard to privacy and confidentiality. Simply anonimizing the data is not enough to fulfil these requirements [@ohm_broken_2009]. Over the years, several other techniques have been used to increase the confidentiality of the data, such as categorizing continuous variables, top coding values above a certain threshold or adding random noise [@drechsler_synthetic_2011]. However, these methods may distort relationships between variables, thereby drastically reducing the data quality and the usability for further research.

@@

The `mice` package has been developed originally for multiple imputation of missing data. In this context, the aim is to replace missing values due to nonresponse by plausible values from the posterior predictive distribution of the variable containing the missing values. Doing so, `mice` makes use of fully conditional specification (FCS), which breaks down the multivariate distribution of the data $\textbf{Y} = (\textbf{Y}_{obs}, \textbf{Y}_{mis})$ into $j = 1, 2, \dots, k$ univariate conditional densities, with $k$ the number of variables in the data. Using FCS, for every variable containing missing values a model is constructed, to impute the missing values $Y_{j, mis}$ with draws from the posterior predictive distribution of $P(Y_{j, mis} | \textbf{Y}_{obs}, \theta)$ on a variable by variable basis. Note that the predictor matrix $Y_{-j}$ may contain yet imputed values from an earlier imputation step, and thus will be updated after every iteration. This procedure is applied $m$ times, resulting in $m$ completed datasets $\textbf{D} = (\textbf{D}^{(1)}, \textbf{D}^{(2)}, \dots, \textbf{D}^{(m)})$, with $\textbf{D}^{(l)} = (\textbf{Y}_{obs}, Y^{(l)}_{mis}$. __MAYBE ADD AN EXAMPLE OF HOW TO DO IT IN `mice`__

__DISCUSS ESTIMATORS FOR MISSING DATA HERE, OR AT A LATER POINT__

This approach can be straightforwardly extended to a synthetic data framework. However, rather than imputing missing data, we overimpute actually observed data, that is, we replace observed values by multiple imputations. For simplicity, assume that the data is completely observed (i.e., $\textbf{Y} = \textbf{Y}_{obs}$). Following the notation of @reiter_raghunathan_multiple_2007, let $Z_j = 1$ if any of the values of unit $j$ are to be replaced by imputations, and $Z_j = 0$ otherwise, with $\textbf{Z} = (Z_1, Z_2, \dots, Z_n)$. Accordingly, the data consists of values that are to be replaced and values that are to be kept (i.e., $\textbf{Y}^{(l)} = (\textbf{Y}^{(l)}_{rep}, \textbf{Y}_{nrep})$. Note the superscript $^{(l)}$ indicating the $l$th synthetic data set. Now, instead of imputing $\textbf{Y}^{(l)}_{mis}$ with draws from the posterior predictive distribution of $P(Y_{j, mis} | \textbf{Y}_{obs}, \theta)$ as in the missing data case, we impute $\textbf{Y}^{(l)}_{rep}$ from the posterior distribution of $(Y_{rep} | \textbf{Y}, \textbf{Z}, \theta)$. This process results in $\textbf{D} = (\textbf{D}^{(1)}, \textbf{D}^{(2)}, \dots, \textbf{D}^{(m)})$, with $m$ the number of synthetic datasets.

By default, `mice` imputes all values that are coded as missing in the dataset. Using the `where` parameter within the mice function, one can choose to impute any value in the dataset. As input, the `where` parameter requires a matrix of the same dimensions of the data, (i.e., a $n \times k$ matrix, with $n$ the number of observations in the collected data) containing logical values.
<!--

Now, we aim to generate the replacement values for the actually observed data $Y_{syn}$, instead of values that replace missing values $Y_{imp}$.

Using FCS, for every variable containing missing values, a model is constructed to impute the missing values on a given variable conditional on all other variables in the data, and the posterior distribution of the parameters used in the model at hand, on a variable by variable basis, which yields $P(Y_{mis} | X, \theta)$.

The `mice` package has been developed originally for multiple imputation of missing data, but can be easily extended to the synthetic data framework. It proceeds by drawing missing or synthetic values from the joint distribution of the observed data. This joint density is generally hard or impossible to derive, which can be overcome by employing fully conditional specification (FCS). Consider a $n \times k$ data matrix $Y$, with values to be replaced $Y^{old}$, the values to be kept $Y^{keep}$ and the values that are already replaced $Y^{new}$, under the assumption that there are no missing values in the observed data set. Instead of drawing from the joint distribution, we draw from $k$ univariate conditional densities for each variable. That is, we draw from the distribution $P\big{(}Y_j^{old} | Y^{keep}, Y^{new}, Y^{old}_{-j}, \theta \big{)}$, where the drawn values are conditional upon the values that are to be kept, the synthetic values that are drawn at an earlier iteration and the values that still are to be replaced, with $j = 1, 2, \dots, k$ and $\theta$ indicating the parameter specifying the distribution of $Y_j^{old}$. Thus, the synthetic values are generated variable by variable. This procedure is executed $i = 1, 2, \dots, m$ times, resulting in $\textbf{D} = \{D^{(1)}, \dots, D^{(m)}\}$ synthetic datasets, with $D^{(i)} = (Y^{keep}, Y^{i, new})$.

@@

-->
